---
title: "How granplot works"
output: html_document
---


```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, cache=TRUE)
# Load any R packages you need here
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggplot2)
library(lvplot)
library(ggridges)
library(tsibbledata)
library(purrr)
library(tsibble)
library(gravitas)
```

<!--temporal granularities and why should we care --> 

`granplot`

Usage: recommends statistical distribution plots of bivariate time granularities that aids systematic exploration

granplot(.data, gran1 = NULL, gran2 = NULL, response = NULL, plot_type = NULL, facet_h = 31, ...) 

- gran1: the first granularity acting as the facet variable
- gran2: the granularity which is to be plotted across x-axis
- response: the univariate time series to be plotted
- plot_type: the type of distribution plot preferred
- facet_h: the highest number of levels for facet variables above which facetting is not recommended 


granplot is useful for exploring statistical distributions for bivariate granularities. It provides the plot recommendations given two granularities. Recommendations for plotting will vary depending on which granularity is placed on the x-axis and which one across facets. 

granplot uses `gran_advice` to obtain the recommendations of distribution plots depending on the levels(very high/high/medium/low) of the two granularities plotted. Assumptions are made to ensure display is not too cluttered by the space occupied by various kinds of distribution plots.

  - very high (facet) and any levels of x-axis

    plots_list <- c("decile", "percentile")
 
  - high (facet) and very high(x-axis)
   
    plots_list <- c("decile", "percentile")
 
  - high (facet) and high(x-axis)

    plots_list <- c("decile", "percentile")
   
  - high(facet) and medium(x-axis)
  
    plots_list <- c("decile", "percentile")   

  - high(facet), low(x-axis)
  
    plots_list <- c("ridge", "violin", "lv", "density")
  
  - medium(facet), very high (x-axis)

    plots_list <- c("decile", "percentile") 
   
  - medium(facet), high (x-axis)
  
    plots_list <- c("decile", "percentile")
  }
  
  - medium(facet), medium (x-axis)
  
    plots_list <- c("decile", "percentile")
 
  
   - medium(facet), low (x-axis)
   
    plots_list <- c("ridge", "violin", "lv", "density")
    
  - low(facet), very high (x-axis)
  
    plots_list <- c("boxplot", "lv", "percentile", "decile")
  
  - low(facet), medium (x-axis)
 
    plots_list <- c("ridge", "violin", "lv", "density", "percentile", "decile")
    
   - low(facet), low (x-axis)

     plots_list <- c("ridge", "violin", "lv", "boxplot", density", "percentile", "decile")
 


# Default values


Default levels are chosen as very high/high/medium/low based on levels of common temporal granularities like day of the month, day of a fortnight or day of a week. Users are free to change these values.

levels_high = 31  
levels_medium = 14  
levels_low = 7  


  very high =  levels <= levels_high  
  high  =   levels_medium < levels < levels_high   
  medium =   levels_low < levels <= levels_medium    
  low  = levels < levels_low  
  
Minimum number of observations for combinations of time granularities  
Assumptions are made to ensure decile/percentile/density based plots are avoided for too few observations


It uses gran_advise to choose the recommended plots if plot_type = NULL.
Now the user might want to plot two temporal granularities that are clashes, they are free to do so however, they will be shown a warning that they are plotting clashes. Warnings will also be shown if number of levels of facet variable too high (inappropriate for facetting), if number of observations are homogenous inter and intra facets. 


However, warnings are generated in the following order:

- If two granularities fed are clashes, then no other warnings will be generated other than that they are clashes.

- If two granularities fed are harmonies, but levels of facets too high for facetting, warnings will be generated.

- If two granularities fed are harmonies and levels of facets not beyond the default or user assigned highest value of  facets allowed, then warnings will be generated in case number of observations for any combination(s) too few to draw a distribution plot or if there is any inter or intra facet heterogeneity.

```{r}
data1 = tsibbledata::nyc_bikes %>% as_tsibble()
data2 = tsibbledata::vic_elec %>% as_tsibble()

gravitas::granplot(data2, gran1="hour_day", gran2 = "month_year", response = "Demand")
```

User is free to choose any one of the seven distribution plots ("ridge", "violin", "lv", "boxplot", density", "percentile", "decile") in plot_type.If plot_type is not specified, the first one from the recommended list is picked up. To see the list of recommended distribution plots, use the function `gran_advice`.

```{r, echo=TRUE}
gravitas:::gran_advice(data2, gran1="hour_day", gran2 = "month_year")
```

