---
title: "Chicago Bikedata Analysis"
author: "Sayani Gupta"
date: "06/06/2019"
output: html_document
---

```{r setup, include=FALSE}
library(forecast)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyverse)
library(lvplot)
library(ggridges)
library(tsibble)
library(bikedata)
```

```{r dataextract, echo=FALSE}
#extract chicago datas for 2017 and 2018
# chicago <- dl_bikedata (city = 'divvy', dates = 2017:2018, data_dir = "bikedata/Data", quiet = TRUE)

# 2018 data doesn't exist

# nyc citibikes for 2017 and 2018 downloaded
#nyc <- dl_bikedata (city = 'citibike', dates = 2017:2018, data_dir = "bikedata/Data", quiet = TRUE)

# 48 monthly files downloaded

file_list <- list.files("bikedata/Data")

for (file in file_list){
       
  # if the merged dataset doesn't exist, create it
  if (!exists("dataset")){
    dataset <- read.table(file, header=TRUE, sep="\t")
  }
   
  # if the merged dataset does exist, append to it
  if (exists("dataset")){
    temp_dataset <-read.table(file, header=TRUE, sep="\t")
    dataset<-rbind(dataset, temp_dataset)
    rm(temp_dataset)
  }
 
}

```

#Data Description

1. Trip Duration (seconds) — How long a trip lasted
2. Start Time and Date - time and date where trip started
3. Stop Time and Date -  time and date where trip ended
4. Start/End Station ID  -  Unique identifier for starting/end bike station
5. Start/End Station Name - Brief address for starting/end bike station
6. Start Station Latitude/Longitude - Coordinates
7. Bike ID - unique identifier for each bike
8. User Type (Customer = 24-hour pass or 3-day pass user;  Subscriber = Annual Member) - Customers are usually tourists, subscribers are usually NYC residents
9. Year of Birth
10. Gender (Zero=unknown; 1=male; 2=female) 


```{r overall_view, echo=FALSE}

nyc_bikes <- tsibbledata::nyc_bikes %>% as_tsibble()
summary(nyc_bikes)

# Top 20 Bike stations by number of starts

top_bikes <- nyc_bikes %>%as_tibble() %>% group_by(start_station) %>% summarise(count = n()) %>% arrange(-count)

top_bikes %>%ggplot(aes(x = reorder(start_station, -count), y= count)) + geom_bar(stat="identity")


nyc_bikes %>%
  as_tibble() %>% 
  dplyr::select(start_station, start_lat, start_long)%>%
  right_join(top_bikes) %>% distinct()

ggplot(nyc_bikes) + geom_polygon(aes(start_long, start_lat, group = group,fill=p1)) + coord_fixed()
```


```{r popular_bikes}

```


```{r busiest_bike_behavior}

```

```{r}

```

