---
title: "Chicago Bikedata Analysis"
author: "Sayani Gupta"
date: "06/06/2019"
output: html_document
---

```{r setup, include=FALSE}
library(forecast)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyverse)
library(lvplot)
library(ggridges)
library(tsibble)
library(bikedata)
library(gravitas)
```

```{r dataextract, echo=FALSE}
#extract chicago datas for 2017 and 2018
# chicago <- dl_bikedata (city = 'divvy', dates = 2017:2018, data_dir = "bikedata/Data", quiet = TRUE)

# 2018 data doesn't exist

# nyc citibikes for 2017 and 2018 downloaded
#nyc <- dl_bikedata (city = 'citibike', dates = 2017:2018, data_dir = "bikedata/Data", quiet = TRUE)

# 48 monthly files downloaded

# file_list <- list.files("bikedata/Data")
# 
# for (file in file_list){
#        
#   # if the merged dataset doesn't exist, create it
#   if (!exists("dataset")){
#     dataset <- read.table(file, header=TRUE, sep="\t")
#   }
#    
#   # if the merged dataset does exist, append to it
#   if (exists("dataset")){
#     temp_dataset <-read.table(file, header=TRUE, sep="\t")
#     dataset<-rbind(dataset, temp_dataset)
#     rm(temp_dataset)
#   }
#  
# }


```

#Data Description

1. Trip Duration (seconds) — How long a trip lasted
2. Start Time and Date - time and date where trip started
3. Stop Time and Date -  time and date where trip ended
4. Start/End Station ID  -  Unique identifier for starting/end bike station
5. Start/End Station Name - Brief address for starting/end bike station
6. Start Station Latitude/Longitude - Coordinates
7. Bike ID - unique identifier for each bike
8. User Type (Customer = 24-hour pass or 3-day pass user;  Subscriber = Annual Member) - Customers are usually tourists, subscribers are usually NYC residents
9. Year of Birth
10. Gender (Zero=unknown; 1=male; 2=female) 


```{r overall_view, echo=FALSE}

nyc_bikes <- tsibbledata::nyc_bikes %>% as_tsibble()
summary(nyc_bikes)

# Top 20 Bike stations by number of starts

top_bikes <- nyc_bikes %>%as_tibble() %>% group_by(start_station) %>% summarise(count = n()) %>% arrange(-count)

#top_bikes %>%ggplot(aes(x = reorder(start_station, -count), y= count)) + geom_bar(stat="identity")


nyc_all <- nyc_bikes %>%
  as_tibble() %>% 
  dplyr::select(start_station, start_lat, start_long)%>%
  right_join(top_bikes) %>% distinct()

ggplot(nyc_all) + geom_polygon(aes(start_long, start_lat,fill=count)) + scale_fill_continuous(type="gradient") + 
  coord_map()
```


```{r popular_bikes}
top_bikes <- nyc_bikes %>%as_tibble() %>% group_by(start_station) %>% summarise(count = n()) %>% arrange(-count) %>% slice(1:20)

top_bikes %>%ggplot(aes(x = reorder(start_station, -count), y= count)) + geom_bar(stat="identity")

```


```{r busiest_bike_behavior}
rides_busiest <- nyc_bikes %>% filter(start_station %in% 3186|end_station %in% 3186)
summary(rides_busiest)

rides_busiest_grans <- rides_busiest  %>% mutate(hour_day = nest("hour", "day", start_time),
                     month_year =   nest("month", "year", start_time),
                     day_week   = nest("day", "week", start_time, abbr),
                     hhour_day   = nest("hhour", "day", start_time),
                     minute_hour = nest("minute", "hour", start_time))

rides_busiest_grans %>% 
  as_tibble %>% group_by(hour_day) %>%
  summarise(Count = n()) %>% 
  ggplot(aes(x = hour_day, y= Count)) + geom_col() + scale_x_continuous(breaks = seq(0, 23, 1))
#8am and 6pm experiences maximum rush in terms of rides

rides_busiest_grans %>% 
  as_tibble %>% group_by(day_week) %>%
  summarise(Count = n()) %>% 
  ggplot(aes(x = day_week, y= Count)) + geom_col() + scale_x_continuous(breaks = seq(1, 7, 1))
#Monday has lowest count when compared with other days of the week

rides_busiest_grans %>% 
  as_tibble %>% group_by(minute_hour) %>%
  summarise(Count = n()) %>% 
  ggplot(aes(x = minute_hour, y= Count)) + geom_line() 


rides_busiest_grans %>% 
  as_tibble %>% group_by(month_year) %>%
  summarise(Count = n()) %>% 
  ggplot(aes(x = month_year, y= Count)) + geom_line() + scale_x_discrete(breaks = seq(1,12,1))
# Mostly in summer months counts are more

granplot = function(data, x, y) {
    data %>% as_tibble %>%  group_by(!!sym(x), !!sym(y)) %>% 
    summarise(ecount = n()) %>% 
     ggplot(aes(x = !!sym(x), y = ecount)) +
          geom_line() +
          theme_bw() + 
          facet_wrap(as.formula(paste("~", y))) + 
          ylab("Total Count") + 
          ggtitle(paste0("Plot of ", x, " given ", y))
}


granplot(rides_busiest_grans, "hhour_day",  "day_week")
granplot(rides_busiest_grans, "hour_day",  "day_week")
granplot(rides_busiest_grans, "day_week",  "month_year")

# Peaks are at different times of the day on Sunday, also there are more than 2 peaks
granplot(rides_busiest_grans, "hour_day",  "day_week")


```

```{r}

```

