---
title: "Exploration of example data"
author: "Sayani Gupta"
date: "27/05/2019"
output: html_document
---

```{r setup, include=FALSE}
library(forecast)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyverse)
library(lvplot)
library(ggridges)
library(tsibble)
```

- **Frequency** ~ 30 minutes interval smart meter data  
- **Time frame** ~ January, 2012 to December 31, 2014 
- **Spread** ~ Australian State (Victoria)


```{r allFig, fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, fig.cap="Various probability distribution plots of electricity consumption data of Victoria from 2012 to 2014. (a) Letter value plot by DoM and MoY, (b) Decile plot by HoD and DoW (c) Box plot by DoY and DoM, (d) Violin plot of DoM and WoM, (e) Ridge plot by WoM and WoY, (f) Decile plot by DoY and MoY. Only plots (a) and (b) show harmonised time variables.", fig.show = 'hold'}

VIC <- read.csv("VIC2015/Data/demand_VIC.csv")
VIC$Date <- as.Date(VIC$Date,origin = "1899-12-30")

first_day_of_month_wday <- function(dx) {
  day(dx) <- 1
  wday(dx)
}

VIC <- VIC %>%mutate(Indx_Year = year(Date),
                     Indx_Month = month(Date, label = FALSE, abbr = TRUE),
                     Indx_Wk_Yr = week(Date),
                     Indx_Wk_Month = ceiling((day(Date) + first_day_of_month_wday(Date) - 1) / 7),
                     Indx_Day_Week = wday(Date, label=FALSE, abbr=TRUE,
                                          week_start=1),

                     Indx_Day_Month = day(Date),

                     Indx_Day_Year =  yday(Date),

                     Indx_Weekend=if_else(Indx_Day_Week %in% c(6,7),1,0),

                     Indx_HlHr_day = Period,
                     month = month(Date, label = FALSE, abbr = TRUE),
         year =  year(Date),
         yday =yday(Date),
         wday = wday(Date, label=FALSE, abbr=TRUE,
                     week_start=1),
         bow = (wday - 1) * 48 + Period,
         dom = day(Date),
         bom = (dom - 1) * 48 + Period,
         Weekend=if_else(wday %in% c(6,7),1,0),
         Indx_hour = ceiling(Period/2),
         Indx_Hour_Yr = Indx_hour + 24*(yday-1),
         Indx_Hour_Month = Indx_hour + 24*(Indx_Day_Month-1),
         Indx_Hour_Wk = Indx_hour + 24*(wday-1))
         
         


VIC <- as_tibble(VIC)



par(mfrow = c(3, 2))

VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Month %in% c(1,2,4,12))%>% ggplot(aes(Indx_Day_Month,OperationalLessIndustrial,group = Indx_Day_Month)) + geom_lv(aes(fill=..LV..),outlier.colour = "red",outlier.shape = 1) + scale_fill_brewer()+ facet_wrap(~Indx_Month) + ylab("Electricity Demand [KWh]") + xlab("Days of the Month") + scale_x_continuous(breaks=seq(0,31,5)) + ggtitle("(a) Letter value plot by DoM and MoY")



VIC_hod_dow <- VIC%>% filter(year %in% c(2012, 2013, 2014)) %>%
  group_by(wday,Indx_hour) %>%
  do({x <- .$OperationalLessIndustrial
  map_dfr(.x = seq(0.1,0.9,0.1),
          .f = ~ tibble(Quantile = .x,
                            Value = quantile(x, probs = .x,na.rm=TRUE)))
  }) %>%  filter(wday %in% c(1,2,6,7))

VIC_hod_dow %>%ggplot(aes(x=Indx_hour,y=Value,col=as.factor(Quantile))) + geom_line() + facet_wrap(~wday) + scale_x_continuous(breaks=seq(1, 24,5)) + ylab("")  + xlab("Day of the Week") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(b) Decile plot by HoD and DoW ")


VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Day_Month %in% c(1,15,29,31))%>% ggplot(aes(as.factor(yday),OperationalLessIndustrial,group = yday)) + geom_boxplot()+ facet_wrap(~Indx_Day_Month) + ylab("Electricity Demand [KWh]") +
  xlab("Days of the Year") + scale_x_discrete(breaks=seq(0,366,60)) +theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Box plot by DoY and DoM")



VIC%>% filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,4))%>% ggplot(aes(as.factor(Indx_Day_Month),OperationalLessIndustrial)) + geom_violin(alpha = 0.03)+ facet_wrap(~Indx_Wk_Month,nrow=3) + ylab("") + xlab("Days of the Month") + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks=seq(0,31,5))+ scale_y_continuous(breaks = seq(2000,9000,2000))+ ggtitle("(d) Violin plot of DoM and WoM")


VIC%>% dplyr:::filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,5),Indx_Wk_Yr <20)%>% ggplot(aes(x=OperationalLessIndustrial,y=as.factor(Indx_Wk_Yr),group=Indx_Wk_Yr)) + geom_density_ridges2() +facet_wrap(~Indx_Wk_Month) + xlab("Electricity Demand [KWh]") + ylab("Weeks of the Year") + scale_x_continuous(breaks = seq(2000,10000,3000)) + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(e) Ridge plot by WoM and WoY")


VIC_moy_doy <- VIC%>% filter(year %in% c(2012, 2013, 2014)) %>%
  group_by(Indx_Month,yday) %>%
  do({x <- .$OperationalLessIndustrial
  map_dfr(.x = seq(0.1,0.9,0.1),
          .f = ~ tibble(Quantile = .x,
                            Value = quantile(x, probs = .x,na.rm=TRUE)))
  }) %>%  filter(Indx_Month %in% c(1,7,11))

VIC_moy_doy %>%ggplot(aes(x=yday,y=Value,col=as.factor(Quantile),group=yday)) + geom_line() + facet_wrap(~Indx_Month)+ scale_x_continuous(breaks=seq(1, 336, 60)) + ylab("")  + xlab("Day of the Year") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot by DoY and MoY.")
```

- **Frequency** ~ 30 minutes interval smart meter data  
- **Time frame** ~ January, 2012 to December 31, 2014 
- **Spread** ~ Australian State (Tasmania)

```{r allFig1s, fig.height=5.5, out.width="50%", fig.pos = "p", fig.align= 'left',echo=FALSE, eval=TRUE, fig.cap="Various probability distribution plots of electricity consumption data of Tasmania from 2012 to 2014. (a)-(b) are examples of Harmonies (c)-(e) are examples of Clashes", fig.show = 'hold'}

TAS <- read.csv("TAS2015/Data/demand_TAS.csv")
TAS$Date <- as.Date(TAS$Date,origin = "1899-12-30")

first_day_of_month_wday <- function(dx) {
  day(dx) <- 1
  wday(dx)
}

TAS <- TAS %>%mutate(Indx_Year = year(Date),
                     Indx_Month = month(Date, label = FALSE, abbr = TRUE),
                     Indx_Wk_Yr = week(Date),
                     Indx_Wk_Month = ceiling((day(Date) + first_day_of_month_wday(Date) - 1) / 7),
                     Indx_Day_Week = wday(Date, label=FALSE, abbr=TRUE,
                                          week_start=1),

                     Indx_Day_Month = day(Date),

                     Indx_Day_Year =  yday(Date),

                     Indx_Weekend=if_else(Indx_Day_Week %in% c(6,7),1,0),

                     Indx_HlHr_day = Period,
                     month = month(Date, label = FALSE, abbr = TRUE),
         year =  year(Date),
         yday =yday(Date),
         wday = wday(Date, label=FALSE, abbr=TRUE,
                     week_start=1),
         bow = (wday - 1) * 48 + Period,
         dom = day(Date),
         bom = (dom - 1) * 48 + Period,
         Weekend=if_else(wday %in% c(6,7),1,0),
         Indx_hour = ceiling(Period/2),
         Indx_Hour_Yr = Indx_hour + 24*(yday-1),
         Indx_Hour_Month = Indx_hour + 24*(Indx_Day_Month-1),
         Indx_Hour_Wk = Indx_hour + 24*(wday-1))
         
         


TAS <- as_tibble(TAS)



par(mfrow = c(3, 2))

TAS%>% filter(year %in% c(2012, 2013, 2014),Indx_Month %in% c(1,2,4,12))%>% ggplot(aes(Indx_Day_Month,OperationalLessIndustrial,group = Indx_Day_Month)) + geom_lv(aes(fill=..LV..),outlier.colour = "red",outlier.shape = 1) + scale_fill_brewer()+ facet_wrap(~Indx_Month) + ylab("Electricity Demand [KWh]") + xlab("Days of the Month") + scale_x_continuous(breaks=seq(0,31,5)) + ggtitle("(a) Letter value plot by DoM and MoY")



TAS_hod_dow <- TAS%>% filter(year %in% c(2012, 2013, 2014)) %>%
  group_by(wday,Indx_hour) %>%
  do({x <- .$OperationalLessIndustrial
  map_dfr(.x = seq(0.1,0.9,0.1),
          .f = ~ tibble(Quantile = .x,
                            Value = quantile(x, probs = .x,na.rm=TRUE)))
  }) %>%  filter(wday %in% c(1,2,6,7))

TAS_hod_dow %>%ggplot(aes(x=Indx_hour,y=Value,col=as.factor(Quantile))) + geom_line() + facet_wrap(~wday) + scale_x_continuous(breaks=seq(1, 24,5)) + ylab("")  + xlab("Day of the Week") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(b) Decile plot by HoD and DoW ")


TAS%>% filter(year %in% c(2012, 2013, 2014),Indx_Day_Month %in% c(1,15,29,31))%>% ggplot(aes(as.factor(yday),OperationalLessIndustrial,group = yday)) + geom_boxplot()+ facet_wrap(~Indx_Day_Month) + ylab("Electricity Demand [KWh]") +
  xlab("Days of the Year") + scale_x_discrete(breaks=seq(0,366,60)) +theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(c) Box plot by DoY and DoM")



TAS%>% filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,4))%>% ggplot(aes(as.factor(Indx_Day_Month),OperationalLessIndustrial)) + geom_violin(alpha = 0.03)+ facet_wrap(~Indx_Wk_Month,nrow=3) + ylab("") + xlab("Days of the Month") + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks=seq(0,31,5))+ scale_y_continuous(breaks = seq(2000,9000,2000))+ ggtitle("(d) Violin plot of DoM and WoM")


TAS%>% dplyr:::filter(year %in% c(2012, 2013, 2014),Indx_Wk_Month %in% c(1,2,5),Indx_Wk_Yr <20)%>% ggplot(aes(x=OperationalLessIndustrial,y=as.factor(Indx_Wk_Yr),group=Indx_Wk_Yr)) + geom_density_ridges2() +facet_wrap(~Indx_Wk_Month) + xlab("Electricity Demand [KWh]") + ylab("Weeks of the Year") + scale_x_continuous(breaks = seq(2000,10000,3000)) + theme(legend.position = "bottom",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(e) Ridge plot by WoM and WoY")


TAS_moy_doy <- TAS%>% filter(year %in% c(2012, 2013, 2014)) %>%
  group_by(Indx_Month,yday) %>%
  do({x <- .$OperationalLessIndustrial
  map_dfr(.x = seq(0.1,0.9,0.1),
          .f = ~ tibble(Quantile = .x,
                            Value = quantile(x, probs = .x,na.rm=TRUE)))
  }) %>%  filter(Indx_Month %in% c(1,7,11))

TAS_moy_doy %>%ggplot(aes(x=yday,y=Value,col=as.factor(Quantile),group=yday)) + geom_line() + facet_wrap(~Indx_Month)+ scale_x_continuous(breaks=seq(1, 336, 60)) + ylab("")  + xlab("Day of the Year") + theme(legend.position = "none",strip.text = element_text(size = 7, margin = margin())) + ggtitle("(f) Decile plot by DoY and MoY.")
```




**Source** ~ Buildings and Property Division at Monash University  **Frequency** ~ 30 minutes interval smart meter data of Monash Residential Services **Time frame** ~ April 4, 2018 to May 31, 2018, **Spread** ~ 73 living units

Living units chosen: B4 29/B3 37

```{r okaygraph1,fig.height=6, fig.width=4, echo=FALSE}

DHR <-read_rds("DHResidence.rds")

selected_units <- DHR %>% filter(Source %in% c("B3 37","B2 15","B3 37", "B4 29", "BG 50"))


selected_units$`Timestamp UTC` <-lubridate::ymd_hms(selected_units$`Timestamp UTC`)

# selected_units_tsibble <- as_tsibble(selected_units,key=id(Source),index=`Timestamp UTC`,tz="UTC")

vic_holidays <- holiday_aus(2018, state = "VIC")

SU_uniform_mutate <-selected_units  %>% mutate(date = date(`Timestamp UTC`),wday = wday(date, label = TRUE, abbr = TRUE,
                       week_start = 1),
                       month = month(date, label = TRUE, abbr = TRUE),
                       year = year(date),
                       hour = hour(`Timestamp UTC`),
                       work = ifelse(wday %in% c("Mon", "Tue", "Wed", "Thu", "Fri"), "Yes", "No"))
                       
                       
# p1 = SU_uniform_mutate %>% filter(Source=="B3 37") %>%
#   dplyr::group_by(date,wday) %>%
#   summarise(dkwh = sum(Value, na.rm=TRUE)) %>%
#   ggplot(aes(x=wday, y=dkwh)) + geom_boxplot() + ylab("Energy consumption in KwH ") +  xlab("Day-of-Week") 
    
p12 = SU_uniform_mutate %>% filter(Source=="B4 29") %>%
  # group_by(date,wday) %>%
  # summarise(dkwh = sum(Value, na.rm=TRUE)) %>%
  ggplot(aes(x=wday, y=Value)) + geom_boxplot() + ylab("Energy consumption in KwH ") + xlab("Day-of-Week") + coord_cartesian(ylim = c(0,0.6))+ ggtitle("a) Across DoY - B4 29") 
p12

p22 = SU_uniform_mutate %>% filter(Source=="B3 37") %>%
  # group_by(date,wday) %>%
  # summarise(dkwh = sum(Value, na.rm=TRUE)) %>%
  ggplot(aes(x=wday, y=Value)) + geom_boxplot() + ylab("Energy consumption in KwH ") + xlab("Day-of-Week") + coord_cartesian(ylim = c(0,0.6))+ ggtitle("b) Across DoW - B3 37") 
p22

p14 = SU_uniform_mutate %>% filter(Source=="B4 29") %>%
  group_by(wday,hour) %>% filter(wday %in% c("Tue","Wed","Thu","Fri","Sat","Sun")) %>% 
  ggplot(aes(x=as.factor(hour), y=Value, group=hour)) + geom_boxplot() +
  facet_wrap(~wday) + 
  ylab("Daily Consumption in KwH") +
  xlab("Hour-of-Day") +  
    coord_cartesian(xlim= c(0,23), ylim = c(0,0.75))+ scale_x_discrete(breaks=c(0,5,10,15,20))+  ggtitle("B4 29")+ ggtitle("c) Across DOW and HoD - B4 29") 
p14


p24 = SU_uniform_mutate %>% filter(Source=="B3 37") %>%
  group_by(wday,hour) %>% filter(wday %in% c("Tue","Wed","Thu","Fri","Sat","Sun")) %>% 
  ggplot(aes(x=as.factor(hour), y=Value, group=hour)) + geom_boxplot() +
  facet_wrap(~wday) + 
  ylab("Daily Consumption in KwH") +
  xlab("Hour-of-Day") +  
    coord_cartesian(xlim= c(0,23), ylim = c(0,0.75))+ scale_x_discrete(breaks=c(0,5,10,15,20))+  ggtitle("B4 29")+ ggtitle("d) Across DoW and HoD - B3 37") 
p24


s1 <- SU_uniform_mutate %>%
  mutate(work = ifelse(date %in% vic_holidays$date, "No", work))


b12_Quantile <- s1 %>% filter(Source=="B4 29") %>% group_by(work,hour) %>% 
   do({x <- .$Value
   map_dfr(.x = seq(0.1,0.9,0.1),
           .f = ~ tibble(Quantile = .x,
                             Value = quantile(x, probs = .x,na.rm=TRUE)))
   })

p16 <- b12_Quantile %>% ggplot(aes(x=hour,y=Value,col=as.factor(Quantile))) +geom_line() + scale_x_continuous(breaks=seq(1, 24, 5)) +theme(legend.position = "bottom") + facet_wrap(~work, labeller = "label_both") + ylab("Daily Consumpttion in KwH") +  xlab("Hour-of-Day") + coord_cartesian(ylim = c(0,0.75)) +  ggtitle("B4 29") + ggtitle("e) Across HoD and WD/NWD - B4 29") 

p16



b22_Quantile <- s1 %>% filter(Source=="B3 37") %>% group_by(work,hour) %>% 
   do({x <- .$Value
   map_dfr(.x = seq(0.1,0.9,0.1),
           .f = ~ tibble(Quantile = .x,
                             Value = quantile(x, probs = .x,na.rm=TRUE)))
   })

p26 <- b22_Quantile %>% ggplot(aes(x=hour,y=Value,col=as.factor(Quantile))) +geom_line() + scale_x_continuous(breaks=seq(1, 24, 5)) +theme(legend.position = "bottom") + facet_wrap(~work, labeller = "label_both") + ylab("Daily Consumpttion in KwH") +  xlab("Hour-of-Day") + coord_cartesian(ylim = c(0,0.75)) +  ggtitle("B4 29") + ggtitle("f) Across HoD and WD/NWD - B3 37") 
p26
```
